---
title: "GSMNP"
author: "Alex Baecher"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r message=FALSE, warning=FALSE}
# load packages

pacman::p_load(
  "colortools",  
  "cowplot",
  "elevatr",
  "ggpubr",
  "ggregplot",
  "ggspatial",
  "ggthemes",
  "INLA",
  "inlabru",
  "INLAutils",
  "lubridate",
  "magick",
  "mapproj",
  "MCMCglmm",
  "osmdata",
  "raster",
  "rayshader",
  "readr",
  "rgdal",
  "rgeos",
  "rgl",
  "scales",
  "sf",
  "shadowtext",
  "sp",
  "spatialEco",
  "spatstat",
  "tidyverse", 
  "tigris",
  "viridis",
  "wesanderson")

```

```{r}
proj <- "+proj=utm +zone=19 +ellps=GRS80 +datum=NAD83"

# setting download method to "libcurl" because raster::getData() and download.file() failed out due to lack of server connection
options(download.file.method = "libcurl")
```

```{r message=FALSE, warning=FALSE}
df <- read_csv("data/SalamanderData_AlexBaecher.csv", na = c(""," ","<NA>","NA",".","..")) %>%
  # Filter out NAs from important columns
  filter(!is.na(Latitude),
         !is.na(Longitude),
         !is.na(EvidenceDate),
         !is.na(StudyId)) %>%
  # Select only columns for study id, localities, and species
  dplyr::select(EvidenceDate,
                Latitude, 
                Longitude, 
                StudyId, 
                Genus, 
                Species) %>%
  # Rename date column
  rename(time = EvidenceDate) %>%
  # Begin manipulating data to consolidate multiple observations into a counts column
  group_by(across(everything())) %>%
  mutate(count = n(),
         time = mdy(time),
         year = year(time),
         decade = format(floor_date(time, years(10)), "%Y"),
         sub_year = as.numeric(substring(year, first = 4))) %>%
  # Filter out observations with problematic date information
  filter(!is.na(decade)) %>%
  ungroup() %>%
  distinct()

df_count_spp <- df %>%
  dplyr::select(-Genus) %>%
  pivot_wider(names_from = Species, values_from = count, values_fill = 0) %>%
  dplyr::select(-time) %>%
  rename("EURWIL" = "Eurycea wilderae", 
         "DESMON" = "Desmognathus monticola", 
         "DESOCO" = "Desmognathus ocoee",
         "PLEJOR" = "Plethodon jordani",
         "DESIMI" = "Desmognathus imitator",
         "DESWRI" = "Desmognathus wrighti") %>%
  mutate(decade = as.factor(decade))

df_count <- df_count_spp %>%
  pivot_longer(cols = c(EURWIL,DESMON,DESOCO,PLEJOR,DESIMI,DESWRI), values_to = "count", names_to = "Species") 

df_count_pjor <- df_count %>%
  filter(Species == "PLEJOR") %>%
  filter(year %in% 1998:2000)

df_R <- df_count_spp %>%
  mutate_if(is.integer, function(x){ifelse(x > 0, 1, 0)}) %>%
  mutate(rich = dplyr::select(., EURWIL:DESWRI) %>% rowSums()) %>%
  dplyr::select(Latitude, Longitude, StudyId, year, decade, rich) 

df_sub_R <- df_R %>%
  filter(year <= 2000)

```

```{r}
# Color palette
pal <- wes_palette("Zissou1", 256, type = "continuous")
```

```{r}
# states <- states(class = "sp") %>%
#   subset(NAME %in% c("Tennessee", "North Carolina", "Mississippi", "Alabama",
#                      "Georgia", "Florida", "South Carolina", "North Carolina")) %>%
#   spTransform(proj)
# 
# write_rds(states, "data/se.rds")
states <- read_rds("data/se.rds")

# Download polygon for TN and NC
tn_nc <- states %>%
  subset(NAME %in% c("Tennessee", "North Carolina"))

# Create polygon for GSMNP
gsmnp <- readOGR("shapefiles/GRSM_BOUNDARY_POLYGON.shp") %>%
  subset(SHAPE_Area > 2e+08) %>%
  spTransform(proj)

# Convert salamander dataframe to spatialpointsdataframe
gsmnp_spdf <- sp::SpatialPointsDataFrame(data = df, 
                                              coords = dplyr::select(df, Longitude, Latitude), 
                                              proj4string = crs("+proj=longlat +datum=NAD83 +no_defs")) %>%
  spTransform(proj) %>%
  erase.point(gsmnp, inside = F)
gsmnp_spdf$Latitude_utm <- gsmnp_spdf@coords[,2]
gsmnp_spdf$Longitude_utm <- gsmnp_spdf@coords[,1]

df_sub_R_spdf <- sp::SpatialPointsDataFrame(data = df_sub_R, 
                                              coords = dplyr::select(df_sub_R, Longitude, Latitude), 
                                              proj4string = crs("+proj=longlat +datum=NAD83 +no_defs")) %>%
  spTransform(proj) %>%
  erase.point(gsmnp, inside = F)
df_sub_R_spdf$Latitude_utm <- df_sub_R_spdf@coords[,2]
df_sub_R_spdf$Longitude_utm <- df_sub_R_spdf@coords[,1]

df_count_pjor_spdf <- sp::SpatialPointsDataFrame(data = df_count_pjor, 
                                              coords = dplyr::select(df_count_pjor, Longitude, Latitude), 
                                              proj4string = crs("+proj=longlat +datum=NAD83 +no_defs")) %>%
  spTransform(proj) %>%
  erase.point(gsmnp, inside = F)
df_count_pjor_spdf$Latitude_utm <- df_count_pjor_spdf@coords[,2]
df_count_pjor_spdf$Longitude_utm <- df_count_pjor_spdf@coords[,1]
```

```{r}
gsmnp_trails_sf <- readOGR("shapefiles/GRSM_TRAILS.shp") %>%
  spTransform(proj) %>%
  gIntersection(gsmnp) %>%
  st_as_sf()
```

```{r}
ext <- extent(gsmnp) %>%
  as("SpatialPolygons") %>%
  gBuffer(width = 9000, byid = T) %>%
  spsample(1000, "regular")
crs(ext) <- crs(tn_nc)

plot(gsmnp, col = "grey80"); points(ext, pch = "+", col = "grey20"); lines(tn_nc, col = "maroon"); box(lwd = 2)
```

```{r}
ext_buff <- extent(-871568.4 + 5000,
                   -764399.9 - 1000,
                   4011977 + 5000,
                   4063126 - 3000) %>%
  as("SpatialPolygons")

# alt_gsmnp_raw <- elevatr::get_elev_raster(ext,
#                                       prj = proj,
#                                       z = 12) %>%
#   crop(ext_buff)
# names(alt_gsmnp_raw) <- "alt"
# write_rds(alt_gsmnp_raw, "alt_gsmnp.rds")

alt_gsmnp <- read_rds("alt_gsmnp.rds") %>%
  aggregate(fact = 2)

hillShade_gsmnp <- hillShade(terrain(alt_gsmnp, opt = "slope"),
                             terrain(alt_gsmnp, opt = "aspect")) %>%
  crop(ext_buff) 

gsmnp_ext <- tn_nc %>%
  subset(NAME == "Tennessee") %>%
  gIntersection(ext_buff)

plot(alt_gsmnp); plot(hillShade_gsmnp, alpha = 0.3, col = grey.colors(100), add = T, legend = F); lines(gsmnp); lines(gsmnp_ext, col = "blue")
```

```{r}
# download.file(url = paste('https://databank.illinois.edu/datafiles/k7adc/download'),
#               destfile = "microclimate.zip", mode='wb', cacheOK=FALSE)

gsmnp_centroid <- gCentroid(gsmnp) %>%
  spTransform("+proj=longlat +zone=17 +datum=NAD83 +units=m +no_defs")

gsmnp_wc_proj <- gsmnp %>%
  spTransform("+proj=longlat +datum=WGS84 +no_defs")

tmax <- raster::getData("worldclim",
                        var = "tmax",
                        res = 0.5,
                        lon = gsmnp_centroid@coords[1],
                        lat = gsmnp_centroid@coords[2]) %>%
  mean() %>%
  crop(gsmnp_wc_proj) %>%
  projectRaster(alt_gsmnp) %>%
  resample(alt_gsmnp, method = "bilinear")

prec <- raster::getData("worldclim",
                        var = "prec",
                        res = 0.5,
                        lon = gsmnp_centroid@coords[1],
                        lat = gsmnp_centroid@coords[2]) %>%
  mean() %>%
  crop(gsmnp_wc_proj) %>%
  projectRaster(alt_gsmnp) %>%
  resample(alt_gsmnp, method = "bilinear")


```

```{r}
tn_ext <- tn_nc %>%
  subset(NAME == "Tennessee") %>%
  gIntersection(as(extent(alt_gsmnp), "SpatialPolygons"))

```

```{r}
gsmnp_inset_ext <- ext_buff %>%
  extent() 

states_crop <- states %>%
  extent()
states_crop[3] <- 2810173 

states_cropped <- states %>%
  raster::intersect(states_crop) %>%
  smoothr::smooth("spline")

states_inset <- ggplot(states_cropped) + 
  geom_polygon(aes(x = long, y = lat, group = group), col = "grey40", fill = "white", size = 0.12, alpha = 0.7) +
  annotate(geom = "rect", 
           xmin = gsmnp_inset_ext[1], 
           xmax = gsmnp_inset_ext[2], 
           ymin = gsmnp_inset_ext[3], 
           ymax = gsmnp_inset_ext[4], 
           colour = "maroon", fill = NA, size = 0.5) +
  annotate(geom = "text", label = "Tennessee", x = -856682 - 240000, y = 4053680 + 70000, size = 1.3, fontface = 2, angle = -12) +
  annotate(geom = "text", label = "North Carolina", x = -855000 + 445000, y = 4015190 - 9000, size = 1.3, fontface = 2, angle = -6) +
  theme_void() + 
  coord_quickmap(); states_inset

cols <- complementary(pal[256/3])[2]
```

```{r}
# create extent object using sp object
# gsmnp_ext_obj <- gsmnp %>%
#   spTransform("+proj=longlat +datum=WGS84 +no_defs") %>%
#   extent()
# 
# ## set parameters to gain water body information
# long_range <- c(gsmnp_ext_obj[1],gsmnp_ext_obj[2]); lat_range <- c(gsmnp_ext_obj[3],gsmnp_ext_obj[4])
# osm_bbox <- c(long_range[1],lat_range[1], long_range[2],lat_range[2])
# 
# gsmnp_water_sp <- opq(osm_bbox) %>%
#   add_osm_feature("water") %>%
#   osmdata_sp()
# 
# gsmnp_water <- spTransform(gsmnp_water_sp$osm_polygons, crs(gsmnp)) %>%
#   gIntersection(as(extent(gsmnp), "SpatialPolygons")) %>%
#   st_as_sf()
# write_rds(gsmnp_water, "data/water.rds")
# 
# plot(gsmnp_water, col = "light blue")
# 
# ## download waterways data
# gsmnp_streams_sp <- opq(osm_bbox) %>%
#   add_osm_feature(key = "waterway") %>%
#   osmdata_sp()
# 
# ## transform waterways data into an sf object
# gsmnp_streams <- spTransform(gsmnp_streams_sp$osm_lines, crs(gsmnp)) %>%
#   gIntersection(gsmnp) %>%
#   st_as_sf()
# write_rds(gsmnp_streams, "data/streams.rds")
```

```{r}
gsmnp_water_sf <- read_rds("data/water.rds")
gsmnp_streams_sf <- read_rds("data/streams.rds")

gsmnp_simp <- gsmnp %>%
  gSimplify(1000) %>%
  smoothr::smooth("ksmooth") %>%
  gBuffer(5000, byid = T) %>%
  smoothr::drop_crumbs(units::set_units(10, km^2))
```

```{r}
gsmnp_streams_vect <- as_Spatial(gsmnp_streams_sf) %>%
  terra::vect()
gsmnp_rast <- terra::rast(alt_gsmnp) > Inf

gsmnp_streams_raster <- terra::rasterize(gsmnp_streams_vect, gsmnp_rast) %>%
  terra::distance(grid = T) %>%
  raster()
```

```{r}
# gg_gsmnp <- ggplot() + 
#   geom_raster(data = as.data.frame(alt_gsmnp, xy = T), aes(x = x, y = y, fill = alt)) +
#   geom_raster(data = as.data.frame(hillShade_gsmnp, xy = T), aes(x = x, y = y, alpha = 1 - layer), fill = "gray20") +
#   geom_sf(data = gsmnp_streams_sf, col = "light blue", alpha = 0.3, size = 0.2) + 
#   geom_sf(data = gsmnp_water_sf, col = "steelblue4", alpha = 0.5) +
#   geom_sf(data = gsmnp_trails_sf, col = "grey20", size = 0.5, alpha = 0.5) +
#   geom_polygon(data = gsmnp_ext, aes(x = long, y = lat, group = group), col = cols, fill = NA, size = 0.5, alpha = 0.3) +
#   geom_polygon(data = gsmnp, aes(x = long, y = lat), col = "grey20", fill = NA, size = 0.25, alpha = 0.1) +
#   geom_polygon(data = as(extent(alt_gsmnp), "SpatialPolygons"), aes(x = long, y = lat), col = "black", fill = NA, size = 0.5) +
#   # geom_point(data = as.data.frame(gsmnp_spdf@coords), aes(x = Longitude, y = Latitude), col = "white", shape = 1, size = 1, alpha = 0.3) + 
#   geom_shadowtext(aes(label = "Tennessee"), x = -856682 - 1100, y = 4053680 + 3500, size = 5, colour = "white") +
#   geom_shadowtext(aes(label = "North Carolina"), x = -855000 - 400, y = 4015190 + 5000, size = 5, colour = "white") +
#   ggtitle("Great Smoky Mountains National Park") +
#   scale_fill_gradientn("Elevation (m)", colors = pal, na.value = NA, guide = guide_colorbar(frame.colour = "grey30", ticks.colour = "grey30")) + 
#   scale_alpha(guide = FALSE, range = c(0,1)) + 
#   annotation_north_arrow(height = unit(1.1, "cm"), width = unit(1.1, "cm"), 
#                          pad_x = unit(13.4, "cm"), pad_y = unit(1.0, "cm")) +
#   annotation_scale(width_hint = 0.1, height = unit(0.2, "cm"), pad_x = unit(13.0, "cm"), pad_y = unit(0.6, "cm")) + 
#   theme_map() + 
#   theme(plot.margin = margin(0.5,0,0,0, "cm"),
#         plot.title = element_text(hjust = 0.5, size = 15, face = 2),
#         legend.position = c(0.24, 0.83), 
#         legend.direction = "horizontal",
#         legend.background = element_rect(colour = "black", fill = alpha("white", 0.7)),
#         legend.key = element_rect(colour = "black", size = 4),
#         legend.key.width = unit(0.6, "cm"),
#         legend.key.size = unit(0.3, "cm"),
#         legend.title = element_text(size = 9),
#         legend.text = element_text(size = 5),
#         legend.margin = margin(t = 0.1, b = 0.1, r = 0.3, l = 0.3, unit = "cm")); gg_gsmnp
# 
# gg_comp <- ggdraw() +
#   draw_plot(gg_gsmnp) +
#   draw_plot(states_inset, x = 0.76, y = 0.058, width = 0.25, height = 0.25)
# 
# ggsave("figures/gg_gsmnp_pal.jpeg", width = 8, height = 4, units = "in", dpi = 600)
# 
# gg_data <- gg_comp + 
#   geom_point(data = df_sub_R_spdf@data, 
#                aes(x = Longitude_utm, y = Latitude_utm, size = rich), shape = 1, alpha = 0.6, col = "black")

```

```{r}
# ggplot() +
#   geom_raster(data = as.data.frame(alt_gsmnp_agg, xy = T), aes(x = x, y = y, fill = alt)) +
#   geom_raster(data = as.data.frame(hillShade_gsmnp_agg, xy = T), aes(x = x, y = y, alpha = 1 - layer), fill = "gray20") +
#   geom_point(data = as.data.frame(gsmnp_spdf), aes(x = Longitude_utm, y = Latitude_utm, col = Species), shape = 3, size = 2) +
#   geom_polygon(data = gsmnp_ext, aes(x = long, y = lat, group = group), fill = NA, col = "maroon", size = 0.25) +
#   geom_polygon(data = gsmnp, aes(x = long, y = lat), fill = NA, col = "grey20", size = 0.25) +
#   geom_polygon(data = as(extent(alt_gsmnp), "SpatialPolygons"), aes(x = long, y = lat), fill = NA, col = "black", size = 0.25) +
#   facet_wrap(~year) +
#   scale_fill_gradientn("Elevation (m)", colors = pal, na.value = NA) +
#   scale_alpha(guide = FALSE, range = c(0,1)) +
#   theme_map() +
#   theme(legend.position = "bottom");
# ggsave("figures/gg_gsmnp_pal_species_year.jpeg", width = 12, height = 10, units = "in", dpi = 600)

```

```{r}
cov_stack <- stack(alt_gsmnp,
                   terrain(alt_gsmnp, opt = "slope"),
                   terrain(alt_gsmnp, opt = "aspect"),
                   terrain(alt_gsmnp, opt = "TPI"),
                   terrain(alt_gsmnp, opt = "roughness"),
                   terrain(alt_gsmnp, opt = "flowdir"),
                   gsmnp_streams_raster,
                   prec,
                   tmax) %>%
  mask(gsmnp) %>%
  scale() 

names(cov_stack) <- c("alt", "slope", "aspect", "TPI" , "rough", "flow", "stream", "prec", "tmax")

plot(cov_stack)
```

```{r}
cov_stack_agg <- aggregate(cov_stack, fact = 30)

n <- nlayers(cov_stack_agg)

cov_stack_agg %>%
  stackApply(., rep(1:n, each=3), fun = mean)
```



df_count_pjor_spdf@data <- bind_cols(
  data.frame(Intercept = rep(1, nrow(df_count_pjor_spdf)),
             slope =     raster::extract(cov_stack_agg$slope,  df_count_pjor_spdf),
             aspect =    raster::extract(cov_stack_agg$aspect, df_count_pjor_spdf),
             elevation = raster::extract(cov_stack_agg$alt,    df_count_pjor_spdf),
             tpi =       raster::extract(cov_stack_agg$TPI,    df_count_pjor_spdf),
             rough =     raster::extract(cov_stack_agg$rough,  df_count_pjor_spdf),
             flow =      raster::extract(cov_stack_agg$flow,   df_count_pjor_spdf),
             stream =    raster::extract(cov_stack_agg$stream, df_count_pjor_spdf),
             tmax =      raster::extract(cov_stack_agg$tmax,   df_count_pjor_spdf),
             prec =      raster::extract(cov_stack_agg$prec,   df_count_pjor_spdf)),
  dplyr::select(df_count_pjor_spdf@data,
                c("Latitude",
                  "Longitude",
                  "StudyId",
                  "count"))) %>%
  mutate(Longitude = scale(Longitude),
         Latitude = scale(Latitude))

cov_df <- as.data.frame(xyFromCell(cov_stack_agg, cell = 1:ncell(cov_stack_agg))) %>%
  mutate(slope =     values(cov_stack_agg$slope),
         aspect =    values(cov_stack_agg$aspect),
         elevation = values(cov_stack_agg$alt),
         tpi =       values(cov_stack_agg$TPI),
         rough =     values(cov_stack_agg$rough),
         flow =      values(cov_stack_agg$flow),
         stream =    values(cov_stack_agg$stream),
         tmax =      values(cov_stack_agg$tmax),
         prec =      values(cov_stack_agg$prec)) %>%
  as.data.frame()

ggplot(cov_df) + 
  geom_point(aes(x = x, y = y, col = aspect), size = 1.3) 

```

```{r}
df_count_pjor_spdf <- df_count_pjor_spdf %>%
  erase.point(gsmnp_simp, inside = F)

df_count_pjor_spdf_mesh <- inla.mesh.2d(boundary = list(gsmnp_simp), 
             loc = df_count_pjor_spdf@coords, 
             cutoff = 1000,
             max.edge = c(2000, Inf)) # 1000, 20000 (for hi res preds)

jpeg("figures/mesh_count.jpeg", units = "in", width = 7, height = 6, res = 600)
plot(df_count_pjor_spdf_mesh); points(df_count_pjor_spdf, col = "red")
dev.off()

```

```{r}
A_count_pred <- inla.spde.make.A(df_count_pjor_spdf_mesh, loc = as.matrix(cov_df[c("x","y")]))
A_count_est <- inla.spde.make.A(df_count_pjor_spdf_mesh, loc = df_count_pjor_spdf@coords)

count_spde <- inla.spde2.matern(mesh = df_count_pjor_spdf_mesh)
s_ind_count <- inla.spde.make.index(name = "sf2", n.spde = count_spde$n.spde)

obs_stack_count <- inla.stack(data = list(y = df_count_pjor_spdf$count),
                    A = list(A_count_est,
                             1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                    effects = list(sf2 = s_ind_count,
                                   Intercept = rep(1, nrow(df_count_pjor_spdf)),
                                   slope = df_count_pjor_spdf@data$slope,
                                   aspect = df_count_pjor_spdf@data$aspect,
                                   tpi = df_count_pjor_spdf@data$tpi,
                                   rough = df_count_pjor_spdf@data$rough,
                                   flow = df_count_pjor_spdf@data$flow,
                                   stream = df_count_pjor_spdf@data$stream,
                                   elev = df_count_pjor_spdf@data$elevation,
                                   prec = df_count_pjor_spdf@data$prec,
                                   tmax = df_count_pjor_spdf@data$tmax),
                    tag = "Obs")

pred_stack_count <- inla.stack(data = list(y = NA),
                    A = list(A_count_pred,
                             1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                    effects = list(sf2 = s_ind_count,
                                   Intercept = rep(1, nrow(cov_df)),
                                   slope = cov_df$slope,
                                   aspect = cov_df$aspect,
                                   tpi = cov_df$tpi,
                                   rough = cov_df$rough,
                                   flow = cov_df$flow,
                                   stream = cov_df$stream,
                                   elev = cov_df$elevation,
                                   prec = cov_df$prec,
                                   tmax = cov_df$tmax),
                    tag = "Pred")

est_stack_count <- inla.stack(data = list(y = NA),
                    A = list(A_count_est,
                             1, 1, 1, 1, 1, 1, 1, 1, 1, 1),
                    effects = list(sf2 = s_ind_count,
                                   Intercept = rep(1, nrow(df_count_pjor_spdf)),
                                   slope = df_count_pjor_spdf@data$slope,
                                   aspect = df_count_pjor_spdf@data$aspect,
                                   tpi = df_count_pjor_spdf@data$tpi,
                                   rough = df_count_pjor_spdf@data$rough,
                                   flow = df_count_pjor_spdf@data$flow,
                                   stream = df_count_pjor_spdf@data$stream,
                                   elev = df_count_pjor_spdf@data$elevation,
                                   prec = df_count_pjor_spdf@data$prec,
                                   tmax = df_count_pjor_spdf@data$tmax),
                    tag = "Est")

fit_stack_count <- inla.stack(obs_stack_count, est_stack_count)
all_stack_count <- inla.stack(obs_stack_count, pred_stack_count, est_stack_count)

```

```{r}

# formula <- y ~ Trt + Age + V4 + f(Ind, model="iid") + f(rand,model="iid")
# 
# formula <- makeGAM('Age', invariant = '', linear = c('Age', 'Trt', 'V4'), returnstring = FALSE)
# 
# result <- inla(formula, family="poisson", data = Epil, 
#                control.predictor = list(compute = TRUE, link = 1))
#  ggplot_inla_residuals(result, observed, binwidth = 0.2)
#  autoplot(result)
# 
# INLAstep_mods <- INLAstep(fam1 = "poisson", 
#                           df_count_pjor_spdf@data,
#                           in_stack = fit_stack_count,
#                           invariant = "-1 + Intercept + f(sf2, model = count_spde)", # figure out how to include intercept as an invariable parameter (maybe remove it from "dataf")
#                           direction = "backwards",
#                           include = 2:10,
#                           y = "y",
#                           y2 = "count",
#                           num.threads = 4,
#                           powerl = 1,
#                           inter = 1,
#                           thresh = 2)
# 
# INLAstep_mods$best_model$summary.random$sf2[, "mean"]
# 
# best_formula <- as.formula(INLAstep_mods$best_formula)

# -1 + Intercept + aspect + elev + tpi + flow + stream + tmax + prec + f(sf2, model = count_spde)
mod <- inla(y ~ -1 + Intercept + f(sf2, model = count_spde), 
            data = inla.stack.data(all_stack_count),
            family = "poisson",
            num.threads = 4,
            control.predictor = list(A = inla.stack.A(all_stack_count), compute = T),
            control.compute = list(cpo = TRUE, dic = TRUE))

autoplot(mod)

```


```{r}
index_pred_pres <- inla.stack.index(all_stack_count, "Pred")$data

cov_stack$pred <- mod$summary.fitted.values[index_pred_pres, "mean"]
cov_stack$sd <- mod$summary.fitted.values[index_pred_pres, "sd"]

# get lower and upper confidence interval
cov_stack$lower_ci <- cov_stack$pred - 1.96*cov_stack$sd
cov_stack$upper_ci <- cov_stack$pred + 1.96*cov_stack$sd

cov_stack_count_rm <- cov_stack %>%
  # disaggregate(fact = 2, method = "bilinear") %>%
  crop(gsmnp) %>%
  mask(gsmnp)

plot(cov_stack$pred)

max_pred <- cellStats(cov_stack_count_rm$pred, "max") %>%
  exp()
max_ci <- cellStats(cov_stack_count_rm$upper_ci, "max") %>%
  exp()

gg_pred_count <- ggplot() +
  geom_raster(data = as.data.frame(cov_stack_count_rm$pred, xy=T), aes(x = x, y = y, fill = exp(pred))) +
  geom_polygon(data = gsmnp, aes(x = long, y = lat), col = "grey40", fill = NA, size = 0.5, alpha = 0.1) +
  scale_fill_gradientn(colors = pal, na.value = NA, "Predicted Prevalence", limits = c(0,max_pred)) +
  theme_map() + theme(legend.position = "bottom")
gg_uncert_count <- ggplot() +
  geom_raster(data = as.data.frame(cov_stack_count_rm$sd, xy=T), aes(x = x, y = y, fill = sd)) +
  geom_polygon(data = gsmnp, aes(x = long, y = lat), col = "grey40", fill = NA, size = 0.5, alpha = 0.1) +
  scale_fill_gradientn(colors = pal, na.value = NA, "Uncertainty") +
  theme_map() + theme(legend.position = "bottom")
gg_lowerCI_count <- ggplot() +
  geom_raster(data = as.data.frame(cov_stack_count_rm$lower_ci, xy=T), aes(x = x, y = y, fill = exp(lower_ci))) +
  geom_polygon(data = gsmnp, aes(x = long, y = lat), col = "grey40", fill = NA, size = 0.5, alpha = 0.1) +
  scale_fill_gradientn(colors = pal, na.value = NA, "Lower 95 CI", limits = c(0,max_ci)) +
  theme_map() + theme(legend.position = "bottom")
gg_upperCI_count <- ggplot() +
  geom_raster(data = as.data.frame(cov_stack_count_rm$upper_ci, xy=T), aes(x = x, y = y, fill = exp(upper_ci))) +
  geom_polygon(data = gsmnp, aes(x = long, y = lat), col = "grey40", fill = NA, size = 0.5, alpha = 0.1) +
  scale_fill_gradientn(colors = pal, na.value = NA, "Upper 95 CI", limits = c(0,max_ci)) +
  theme_map() + theme(legend.position = "bottom")

gg_sec_count <- ggarrange(gg_pred_count, gg_uncert_count, gg_lowerCI_count, gg_upperCI_count, ncol = 2, nrow = 2);gg_sec_count
ggsave("figures/modeled_counts.jpeg", dpi = 600, width = 10, height = 6, units = "in")

# ggarrange(gg_data, gg_sec_count, nrow = 2, heights = c(7, 10))
# ggsave("figures/comp_modeled_counts.jpeg", dpi = 600, width = 10, height = 10, units = "in")
  

```



